Stochastic Variance-Reduced Policy Gradient

### 2018-02-07 svrpg.tex
corrections on images

Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^d\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$. We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $p_{\vtheta}(\tau)$ with $p(\tau|\vtheta)$ for the sake of readability.
The search for a locally optimal policy is performed through gradient ascent, where the policy gradient
%of $J(\vtheta)$ \wrt the policy parameters 
is \cite{sutton2000policy, Peters2008reinf}:
\begin{align} \label{E:policygradient}
        \gradJ{\vtheta} = \EVV[\tau \sim p(\cdot|\vtheta)]{\score{\vtheta}{\tau}\Reward(\tau)}.
        %\int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
\end{align}
Notice that the distribution defining the gradient is induced by the current policy. This aspect introduces a nonstationarity in the sampling process. Since the underlying distribution changes over time, it is necessary to resample at each update or use weighting techniques such as importance sampling.
Here, we consider the \emph{online learning scenario}, where trajectories are sampled by interacting with the environment at each policy change. 
In this setting, stochastic gradient ascent is typically employed.
At each iteration $k >0$, a batch $\mathcal{D}_N^k = \{\tau_i\}_{i=0}^N$ of $N>0$ trajectories is collected using policy $\pi_{\vtheta_k}$.
The policy is then updated as $\vtheta_{k+1}  = \vtheta_k + \alpha\gradApp{\vtheta_k}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of~\eqref{E:policygradient} using $\mathcal{D}_N^k$. The most common policy gradient estimators can be expressed as follows
\begin{align} \label{E:policygradient.estimate}
        \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} g(\tau_i|\vtheta), \quad \tau_i \in \mathcal{D}_N^k,
\end{align}
where $g(\tau_i|\vtheta)$ is an estimate of $\score{\vtheta}{\tau_i}\Reward(\tau_i)$, \eg REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
Although REINFORCE definition is simpler than the G(PO)MDP one, the latter is often preferred due to its lower variance.
We refer the reader to Appendix~\ref{A:gradient_estimators} for details and a formal definition of $g$.

The main limitation of plain policy gradient is the high variance of these estimators.
The na\"ive approach of increasing the batch size is not an option in RL due to the high cost of collecting samples, \ie by interacting with the environment.
For this reason, literature has focused on the introduction of baselines aiming to reduce the variance~\citep[\eg][]{Peters2008reinf,Thomas2017actionbaseline,wu2018variance}.\footnote{A baseline is a function $f(s,a)$ such that $\EVV[a \sim \pol]{f(s,a)} = 0$, for any $s\in\Sspace$. \todopir{check the definition of baseline}}
There has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
Although these techniques have been mainly derived for finite-sum problems, we will show in Sec.~\ref{sec:alg} how they can be used in RL.
The next section has the aim to describe variance reduction techniques for finite-sum problems. In particular, we will present the SVRG algorithm that is at the core of this work.

@@

\section{SVRG in Reinforcement Learning}\label{sec:alg}
In online RL problems the usual approach is to tune the batch size of SG in order to find the optimal trade-off between variance and speed.
Recall that compared to SL, the samples are not fixed in advance but need to be collected at each policy change.
Since this operation may be costly, we would like to minimize the number of interactions with the environment.
Due to this issues, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
However, a direct application of SVRG to RL is not possible due to the following properties:
% In on-line policy gradient problems, SG is not really a choice, but is dictated by the necessity of interacting with an unknown environment.\todopir{I do not understand the meaning of this sentence. I would say that SG is not a choice by referring to papers that uses batches}
% We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence. In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$. Compared to the typical finite sum optimization scenario, we have the following additional challenges:
\begin{description}
        \item[Non-concavity:] the objective function $J(\vtheta)$ is typically non-concave.
        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The actual dataset is infinite and coincide with $\pi_{\vtheta}(\tau)$.
%    \item[Approximation:] the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the parameter $\vtheta$ influences the sampling process.
\end{description}
To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}, see App.~\ref{app:gauss} for the derivation.
Because of the infinite dataset, we can only rely on an estimate of the full gradient.
\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being \textcolor{red}{concave}, showing that SVRG is robust to inexact computation of the full gradient. In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate. {\color{red}Later on we will show how the estimation accuracy impacts on the convergence results with non-concave objective.} 
Finally, the unbiasedness of Eq.~\ref{E:svrg.gradient.correction} is broken by the non-stationarity of the optimization problem.
To overcome this limitation we employ importance weighting~\citep[\eg][]{rubinstein1981simulation,precup2000eligibility} to correct the distribution shift.
%Finally, we employ importance weighting~\citep[\eg]{rubinstein1981simulation,precup2000eligibility} to guarantee proper sampling of trajectories.

We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG) for a generic policy gradient estimator $g$ (see Eq.~\ref{E:policygradient.estimate}).
% For simplicity, we focus on the REINFORCE estimator but the extension to G(PO)MDP is straightforward.
Between snapshots, the SVRPG estimate is corrected using importance weighting, for any $t \in \{0,\ldots,m-1\}$
\begin{align*}
        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
\end{align*}
% \begin{align*}
% 	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\
% 		&+ \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
% 		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i),
% \end{align*}
where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{p(\tau|\wt{\vtheta})}{p(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$. 
Similarly to SVRG, we have that $\vtheta_0 := \wt{\vtheta}$ and the update is a FG step.
Our update is still fundamentally on-policy, since the weighting concerns only the correction term. However, this partial ``off-policyness'' is enough to introduce variance. This is a well-known issue of importance sampling~\citep[\eg][]{thomas2015high}. To mitigate it, we use mini-batches of trajectories of size $B \ll N$ to average the correction, \ie
\begin{align}\label{E:svrpg.estimate.batch}
        \blacktriangledown J(\vtheta_{t}) &:= v_t= \wh{\nabla}_N J(\wt{\vtheta})\\ \notag
                                            & \quad{} + 
        \underbracket{
        \frac{1}{B} \sum_{i=0}^{B-1} \left[
        g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
\right]}_{c_t}.
\end{align}
We want just to stress that $\EVV[\tau_i \sim p(\cdot|\vtheta_t)]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$.\footnote{The reader can refer to App.~\ref{} for off-policy gradients and variants of REINFORCE and G(PO)MDP.}
This property will be used to prove Lem.~\ref{L:svrpg.properties}.
%\begin{align*}
%&\blacktriangledown J(\vtheta) = \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) +\\
%&\quad\frac{1}{B}\left[\sum_{i=0}^{B-1}
%\nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
%- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i)\right].
%\end{align*}
The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}. It is easy to show that the SVRPG estimator has the following, desirable properties (see App.~\ref{app:proofs}):
\begin{lemma}\label{L:svrpg.properties} \todopirout{extended to any unbiased estimator}
        {\color{blue} Let $\wh{\nabla}_N J(\vtheta)$ be an unbiased estimator of~\eqref{E:policygradient} }
%Denote by $\wh{\nabla}_N^{\textsc{RF}} J(\vtheta)$ the REINFORCE estimator in~\eqref{E:policygradient.estimate} 
and let $\vtheta^* \in \argmin_{\vtheta} \{J(\vtheta)\}$. Then, the SVRG estimate in~\eqref{E:svrpg.estimate.batch} is \emph{unbiased}
\[
\mathop{\mathbb{E}}
\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
\]
and regardless of the mini-batch size $B$:\footnote{
For any vector $\mathbf{x}$, we use $\Var[\mathbf{x}]$ to denote the trace of the covariance matrix, \ie $\Tr\EVV[]{(\mathbf{x}-\EVV[]{\mathbf{x}})(\mathbf{x}-\EVV[]{\mathbf{x}})^T}$.}
\[
	\Var\left[\gradBlack{\vtheta^*}\right] = 
    \Var\left[\wh{\nabla}_N J(\vtheta^*)\right].
\]
\end{lemma}

%\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
%The SVRPG estimator is unbiased:
%\[
%\mathop{\mathbb{E}}
%\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
%\]
%\end{restatable}
%\begin{restatable}{lemma}{zerovar}\label{lemma:zerovar}
%At convergence, the variance of the SVRPG estimator is equal to the variance of the REINFORCE estimator, regardless of the mini-batch size $B$:
%\[
%	\mathbb{V}ar\left[\gradBlack{\vtheta^*}\right] = 
%	\mathbb{V}ar\left[\gradApp{\vtheta^*}{N}\right].
%\]
%\end{restatable}
In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can actually achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
The resulting SVRPG policy optimization method is detailed in Alg.~\ref{alg:svrpg}.\footnote{Note that when $t=0$ we can avoid to sample trajectories and instead perform directly the policy update using the FG. For sake of clarity, we have not explicitly differentiate this step from the standard update rule in Alg.~\ref{alg:svrpg}.}

To obtain a method that can be used in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient, the mini-batch size $B$, and the step size $\alpha$.
Adaptive selection of these meta-parameters is also possible, and is explored in Sec.~\ref{sec:stopping}.
However, the next section will provide general converge guarantees for SVRPG \textcolor{red}{with REINFORCE estimator}.

\begin{algorithm}[tb]
        \caption{SVRPG}
	\label{alg:svrpg}
	\begin{algorithmic}
        \STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, batch size $N$, mini-batch size $B$, gradient estimator $g$, initial parameter $\vtheta_{0}^m := \wt{\vtheta}^0$
		\FOR{$s=0$ {\bfseries to} $S-1$}
        \STATE $\vtheta_0^{s+1} := \tilde{\vtheta}^{s} = \vtheta_{m}^s$
        \STATE Sample $N$ trajectories $\{\tau_j\}$ from $p(\cdot\vert\tilde{\vtheta}^{s})$
        \STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\ref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
		\FOR{$t=0$ {\bfseries to} $m-1$}
        \STATE Sample $B$ trajectories $\{\tau_i\}$ from $p(\cdot\vert\vtheta_t^{s+1})$
%        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)$
        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
        \begin{aligned}[t]
                \Big( & g(\tau_i|\vtheta_t^{s+1})\\ 
                      & \quad{} - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \Big)
        \end{aligned}$
%        \STATE $\begin{aligned}[t]
%                &v^{s+1}_t = \wt{\mu}\\ 
%                & +\frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)
%                \end{aligned}$
        \STATE $v^{s+1}_t = \wt{\mu} + c^{s+1}_t$ % \gradApp{\wt{\vtheta}^{s}}{N} + c^{s}_t$
%%		\begin{align*}
%%		\blacktriangledown J(\vtheta_t^{s+1}) = 
%%		&\gradApp{\tilde{\vtheta}^s}{N} \\
%%		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
%%		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
%%		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \tilde{\vtheta}^{s}}\Reward(\tau_i)\right]
%%		\end{align*}
        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
		\ENDFOR
		\ENDFOR
        \STATE {\bfseries return} $\vtheta_t^{s+1}$ with $(s,t)$ picked uniformly at random from $\{[0,S-1]\times[0,m-1]\}$
	\end{algorithmic}
\end{algorithm}

\section{Convergence Guarantees of SVRPG}\label{sec:conv}
In this section, we state the convergence guarantees for SVRPG {\color{red}with REINFORCE estimator}.\footnote{\color{red}Similar results can be derived for G(PO)MDP, see App.~\ref{app:}.}
We mainly leverage on the recent analysis of \textcolor{red}{non-concave} SVRG~\cite{reddi2016stochastic,allen2016variance}.
Each of the three challenges presented at the beginning of Sec.~\ref{sec:alg} can potentially prevent convergence, so we need additional assumptions.
In App.~\ref{app:gauss} we show how the widespread Gaussian policy satisfies these assumptions.

\textit{1) Non-concavity.} A common assumption in this case it to assume the objective function to be $L$-smooth.
However, in RL we can consider the following assumption which is sufficient for the $L$-smoothness of the objective (see Lem.~\ref{lemma:lsmooth}).
	\begin{restatable}[On policy derivatives]{assumption}{boundedscore}\label{ass:bounded_score}
		For each state-action pair $(s,a)$, value of $\vtheta$ and dimensions $i,j$ there are positive constants $G,F<\infty$ such that:
\[
		\left|\nabla_{\theta_i}\log\pi_{\vtheta}(a\vert s)\right| \leq \GRADLOG, \qquad
        \left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(a \vert s)\right| \leq \HESSLOG.
\]
	\end{restatable}

>>>>

Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^d\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$. For sake of brevity, we will denote the performance of a parametric policy with $J(\vtheta)$ and the probability of a trajectory $\tau$ with $p(\tau|\vtheta)$ (in some occasions, $p(\tau|\vtheta)$ will be replaced by $p_{\vtheta}(\tau)$ for the sake of readability).
The search for a locally optimal policy is performed through gradient ascent, where the policy gradient
%of $J(\vtheta)$ \wrt the policy parameters 
is \cite{sutton2000policy, Peters2008reinf}:
\begin{align} \label{E:policygradient}
        \gradJ{\vtheta} = \EVV[\tau \sim p(\cdot|\vtheta)]{\score{\vtheta}{\tau}\Reward(\tau)}.
        %\int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
\end{align}
Notice that the distribution defining the gradient is induced by the current policy. This aspect introduces a nonstationarity in the sampling process. Since the underlying distribution changes over time, it is necessary to resample at each update or use weighting techniques such as importance sampling.
Here, we consider the \emph{online learning scenario}, where trajectories are sampled by interacting with the environment at each policy change. 
In this setting, stochastic gradient ascent is typically employed.
At each iteration $k >0$, a batch $\mathcal{D}_N^k = \{\tau_i\}_{i=0}^N$ of $N>0$ trajectories is collected using policy $\pi_{\vtheta_k}$.
The policy is then updated as $\vtheta_{k+1}  = \vtheta_k + \alpha\gradApp{\vtheta_k}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of Eq.~\eqref{E:policygradient} using $\mathcal{D}_N^k$. The most common policy gradient estimators (\eg REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}) can be expressed as follows
\begin{align} \label{E:policygradient.estimate}
        \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} g(\tau_i|\vtheta), \quad \tau_i \in \mathcal{D}_N^k,
\end{align}
where $g(\tau_i|\vtheta)$ is an estimate of $\score{\vtheta}{\tau_i}\Reward(\tau_i)$.
Although the REINFORCE definition is simpler than the G(PO)MDP one, the latter is usually preferred due to its lower variance.
We refer the reader to Appendix~\ref{A:gradient_estimators} for details and a formal definition of $g$.

The main limitation of plain policy gradient is the high variance of these estimators.
The na\"ive approach of increasing the batch size is not an option in RL due to the high cost of collecting samples, \ie by interacting with the environment.
For this reason, literature has focused on the introduction of baselines aiming to reduce the variance~\citep[\eg][]{Peters2008reinf,Thomas2017actionbaseline,wu2018variance}.\footnote{A baseline is a function $f(s,a)$ such that $\EVV[a \sim \pol]{f(s,a)} = 0$, for any $s\in\Sspace$. \todopir{check the definition of baseline}}
There has been a surge of recent interest in variance reduction techniques for gradient optimization in supervised learning.
Although these techniques have been mainly derived for finite-sum problems, we will show in Section~\ref{sec:alg} how they can be used in RL.
The next section has the aim to describe variance reduction techniques for finite-sum problems. In particular, we will present the SVRG algorithm that is at the core of this work.

@@

\section{SVRG in Reinforcement Learning}\label{sec:alg}
In online RL problems, the usual approach is to tune the batch size of SG to find the optimal trade-off between variance and speed.
Recall that, compared to SL, the samples are not fixed in advance but we need to collect them at each policy change.
Since this operation may be costly, we would like to minimize the number of interactions with the environment.
For these reasons, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
However, a direct application of SVRG to RL is not possible due to the following issues:
% In on-line policy gradient problems, SG is not really a choice, but is dictated by the necessity of interacting with an unknown environment.\todopir{I do not understand the meaning of this sentence. I would say that SG is not a choice by referring to papers that uses batches}
% We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence. In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$. Compared to the typical finite sum optimization scenario, we have the following additional challenges:
\begin{description}
        \item[Non-concavity:] the objective function $J(\vtheta)$ is typically non-concave.
        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The objective function is an expected value over the trajectory density $p_{\vtheta}(\tau)$ of the total discounted reward, for which we would need an infinite dataset.
%    \item[Approximation:] the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the policy parameter $\vtheta$ influences the sampling process.
\end{description}
To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}, see Appendix~\ref{app:gauss} for the derivation.
Because of the infinite dataset, we can only rely on an estimate of the full gradient.
\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being \textcolor{red}{concave}, showing that SVRG is robust to an inexact computation of the full gradient. In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate. {\color{red}Later on we will show how the estimation accuracy impacts on the convergence results with non-concave objective.} 
Finally, the non-stationarity of the optimization problem introduces a bias into the SVRG estimator in Eq.~\eqref{E:svrg.gradient.correction}.
To overcome this limitation we employ importance weighting~\citep[\eg][]{rubinstein1981simulation,precup2000eligibility} to correct the distribution shift.
%Finally, we employ importance weighting~\citep[\eg]{rubinstein1981simulation,precup2000eligibility} to guarantee proper sampling of trajectories.

We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG) for a generic policy gradient estimator $g$ (see Eq.~\eqref{E:policygradient.estimate}).
% For simplicity, we focus on the REINFORCE estimator but the extension to G(PO)MDP is straightforward.
Between snapshots, the SVRPG estimate is corrected using importance weighting, for any $t \in \{0,\ldots,m-1\}$
\begin{align*}
        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta}),
\end{align*}
% \begin{align*}
% 	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\
% 		&+ \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
% 		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i),
% \end{align*}
where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in Eq.~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{p(\tau|\wt{\vtheta})}{p(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$. 
Similarly to SVRG, we have that $\vtheta_0 := \wt{\vtheta}$, and the update is a FG step.
Our update is still fundamentally on-policy since the weighting concerns only the correction term. However, this partial ``off-policyness'' is enough to introduce variance. This is a well-known issue of importance sampling~\citep[\eg][]{thomas2015high}. To mitigate it, we use mini-batches of trajectories of size $B \ll N$ to average the correction, \ie
\begin{align}\label{E:svrpg.estimate.batch}
        \blacktriangledown J(\vtheta_{t}) &:= v_t= \wh{\nabla}_N J(\wt{\vtheta})\\ \notag
                                            & \quad{} + 
        \underbracket{
        \frac{1}{B} \sum_{i=0}^{B-1} \left[
        g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
\right]}_{c_t}.
\end{align}
We want just to stress that $\EVV[\tau_i \sim p(\cdot|\vtheta_t)]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$.\footnote{The reader can refer to Appendix~\ref{} for off-policy gradients and variants of REINFORCE and G(PO)MDP.}
This property will be used to prove Lemma~\ref{L:svrpg.properties}.
%\begin{align*}
%&\blacktriangledown J(\vtheta) = \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) +\\
%&\quad\frac{1}{B}\left[\sum_{i=0}^{B-1}
%\nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
%- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i)\right].
%\end{align*}
The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}. It is easy to show that the SVRPG estimator has the following, desirable properties (see Appendix~\ref{app:proofs}):
\begin{lemma}\label{L:svrpg.properties} \todopirout{extended to any unbiased estimator}
        {\color{blue} Let $\wh{\nabla}_N J(\vtheta)$ be an unbiased estimator of~\eqref{E:policygradient} }
%Denote by $\wh{\nabla}_N^{\textsc{RF}} J(\vtheta)$ the REINFORCE estimator in~\eqref{E:policygradient.estimate} 
and let $\vtheta^* \in \argmin_{\vtheta} \{J(\vtheta)\}$. Then, the SVRG estimate in~\eqref{E:svrpg.estimate.batch} is \emph{unbiased}
\[
\mathop{\mathbb{E}}
\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
\]
and regardless of the mini-batch size $B$:\footnote{
For any vector $\mathbf{x}$, we use $\Var[\mathbf{x}]$ to denote the trace of the covariance matrix, \ie $\Tr\EVV[]{(\mathbf{x}-\EVV[]{\mathbf{x}})(\mathbf{x}-\EVV[]{\mathbf{x}})^T}$.}
\[
	\Var\left[\gradBlack{\vtheta^*}\right] = 
    \Var\left[\wh{\nabla}_N J(\vtheta^*)\right].
\]
\end{lemma}

%\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
%The SVRPG estimator is unbiased:
%\[
%\mathop{\mathbb{E}}
%\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
%\]
%\end{restatable}
%\begin{restatable}{lemma}{zerovar}\label{lemma:zerovar}
%At convergence, the variance of the SVRPG estimator is equal to the variance of the REINFORCE estimator, regardless of the mini-batch size $B$:
%\[
%	\mathbb{V}ar\left[\gradBlack{\vtheta^*}\right] = 
%	\mathbb{V}ar\left[\gradApp{\vtheta^*}{N}\right].
%\]
%\end{restatable}
In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
Algorithm~\ref{alg:svrpg} presents the details of the SVRPG method for policy optimization.\footnote{Note that when $t=0$ we can avoid to sample trajectories and instead directly perform the policy update using the FG. For the sake of clarity, we have not explicitly differentiated this step from the standard update rule in Algorithm~\ref{alg:svrpg}.}

To use SVRPG in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient, the mini-batch size $B$, and the step size $\alpha$.
Adaptive selection of these meta-parameters is also possible, and it is explored in Section~\ref{sec:stopping}.
However, the next section will provide general converge guarantees for SVRPG \textcolor{red}{with REINFORCE estimator}.

\begin{algorithm}[tb]
        \caption{SVRPG}
	\label{alg:svrpg}
	\begin{algorithmic}
        \STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, batch size $N$, mini-batch size $B$, gradient estimator $g$, initial parameter $\vtheta_{0}^m := \wt{\vtheta}^0$
		\FOR{$s=0$ {\bfseries to} $S-1$}
        \STATE $\vtheta_0^{s+1} := \tilde{\vtheta}^{s} = \vtheta_{m}^s$
        \STATE Sample $N$ trajectories $\{\tau_j\}$ from $p(\cdot\vert\tilde{\vtheta}^{s})$
        \STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\eqref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
		\FOR{$t=0$ {\bfseries to} $m-1$}
        \STATE Sample $B$ trajectories $\{\tau_i\}$ from $p(\cdot\vert\vtheta_t^{s+1})$
%        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)$
        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
        \begin{aligned}[t]
                \Big( & g(\tau_i|\vtheta_t^{s+1})\\ 
                      & \quad{} - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \Big)
        \end{aligned}$
%        \STATE $\begin{aligned}[t]
%                &v^{s+1}_t = \wt{\mu}\\ 
%                & +\frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)
%                \end{aligned}$
        \STATE $v^{s+1}_t = \wt{\mu} + c^{s+1}_t$ % \gradApp{\wt{\vtheta}^{s}}{N} + c^{s}_t$
%%		\begin{align*}
%%		\blacktriangledown J(\vtheta_t^{s+1}) = 
%%		&\gradApp{\tilde{\vtheta}^s}{N} \\
%%		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
%%		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
%%		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \tilde{\vtheta}^{s}}\Reward(\tau_i)\right]
%%		\end{align*}
        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
		\ENDFOR
		\ENDFOR
        \STATE {\bfseries return} $\vtheta_t^{s+1}$ with $(s,t)$ picked uniformly at random from $\{[0,S-1]\times[0,m-1]\}$
	\end{algorithmic}
\end{algorithm}

\section{Convergence Guarantees of SVRPG}\label{sec:conv}
In this section, we state the convergence guarantees for SVRPG {\color{red}with REINFORCE estimator}.\footnote{\color{red}Similar results can be derived for G(PO)MDP, see Appendix~\ref{app:}.}
We mainly leverage on the recent analysis of \textcolor{red}{non-concave} SVRG~\cite{reddi2016stochastic,allen2016variance}.
Each of the three challenges presented at the beginning of Section~\ref{sec:alg} can potentially prevent convergence, so we need additional assumptions.
In Appendix~\ref{app:gauss} we show how the traditional Gaussian policy satisfies these assumptions.

\textit{1) Non-concavity.} A common assumption, in this case, is to assume the objective function to be $L$-smooth.
However, in RL we can consider the following assumption which is sufficient for the $L$-smoothness of the objective (see Lemma~\ref{lemma:lsmooth}).
	\begin{restatable}[On policy derivatives]{assumption}{boundedscore}\label{ass:bounded_score}
		For each state-action pair $(s,a)$, value of $\vtheta$ and dimensions $i,j$ there are positive constants $G,F<\infty$ such that:
\[
		\left|\nabla_{\theta_i}\log\pi_{\vtheta}(a\vert s)\right| \leq \GRADLOG, \qquad
        \left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(a \vert s)\right| \leq \HESSLOG.
\]
	\end{restatable}
